---
title: "Extended Data Fig. 3: Mutation-cluster examples"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](ExtendedDataFig3.R)

## Overview

The panels illustrate mutation clustering within individual tumors and the relationship between mutation-level allele fractions, estimated cancer cell fractions, and cluster assignments.

## Analyses

### a–b. Within-tumor mutation clusters

DPClust assigns mutations to populations with different inferred cellular fractions. Sample-level cluster visualizations require the fitted clustering output and its plotting inputs.

The plotting implementation for these example panels is not included. The available mutation-assignment table is insufficient to regenerate fitted density displays without the corresponding model output.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `DP_info_data.RData` | Mutation-level values and assigned clusters available to other PTC analyses. |
| DPClust model and plotting inputs | Required for the example density displays; not included in this module. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md).

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" ExtendedDataFig3
```

The figure entry point does not currently produce plots.

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_ExtendedDataFig3(input_dir, output_dir)`.

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/ExtendedDataFig3.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('No local plot implementation is available for this figure.\n')
}
```
