---
title: "Extended Data Fig. 2: Driver associations with CCF-based heterogeneity"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](ExtendedDataFig2.R)

## Overview

Cohort-specific analyses compare CCF-based MATH scores between tumors with and without selected driver alterations.

## Analyses

### MATH definition

Cancer cell fraction (CCF) values are summarized within each tumor as `100 × mad(CCF) / median(CCF)`. R's `mad()` uses its default normal-consistency scaling constant. This analysis uses CCF-based MATH; a separately calculated VAF-based summary is not the plotted outcome.

### a. Driver-association summary

For each eligible alteration category and cohort, a Wilcoxon test compares MATH scores between alteration-positive and alteration-negative tumors. The horizontal axis shows the log2 ratio of group medians; the vertical axis shows the unadjusted P value. Holm-adjusted P values are also calculated within each cohort.

### b. Alteration-stratified distributions

Alteration categories with a nominal P value below 0.05 in any cohort are selected for violin and box plots. Distributions are separated by alteration presence and cohort. Selection uses nominal significance, so these plots are descriptive follow-ups to the association tests.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData` | Cohort membership and colors. |
| `DP_info_data.RData` | Mutation-level CCF and VAF values. |
| `Genome_landscape_manual_final.RData` | Curated driver categories and pooled fusion categories. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md). Use an output directory for your local results. Available plots are exported as PDF and PNG with Roboto Condensed typography.

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" ExtendedDataFig2
```

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_ExtendedDataFig2(input_dir, output_dir)`.

Shared analysis functions: [evolution_helpers.R](../functions/evolution_helpers.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/ExtendedDataFig2.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('## Local figures\n\n')
  files <- file.path(params$figure_dir, c("ExtendedDataFig2a.png","ExtendedDataFig2b.png"))
  for (p in files) {
    if (file.exists(p)) {
      cat(sprintf('<img alt="Extended Data Fig. 2: Driver associations with CCF-based heterogeneity" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    } else {
      cat('Local plot not found: ', basename(p), '\n\n', sep = '')
    }
  }
}
```
