---
title: "Extended Data Fig. 4: BRAF, TERT, and tumor timing"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](ExtendedDataFig4.R)

## Overview

TCGA-THCA tumors are grouped by BRAF and TERT alteration presence to compare estimated MRCA age, diagnosis age, and APOBEC-associated mutagenesis.

## Analyses

### a. Estimated MRCA age

Box plots compare estimated age at the most recent common ancestor (MRCA) across BRAF-only, TERT-only, BRAF-plus-TERT, and remaining tumors. Timing estimates use the 1× acceleration model. Groups are ordered by median MRCA age.

### b. Age at diagnosis

The same alteration groups are compared for diagnosis age and ordered by their median diagnosis age. Panels a and b use two-sided pairwise Wilcoxon tests with Benjamini–Hochberg adjustment.

### c. APOBEC-associated mutagenesis

APOBEC activity is defined by combined SBS2 and SBS13 contributions exceeding 50 mutations and exceeding 5% of the signature burden. Stacked bars show the proportion meeting both criteria within each alteration group.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData`, `wgs_covdata_tp.RData`, `wgs_clinical.RData` | Cohort and clinical data used by the common input loader. |
| `Genome_landscape_manual_final.RData` | BRAF and TERT alteration groups. |
| `thyroid_evolution_analysis/Chronological_timing_short.RData` | Estimated MRCA age. |
| `thyroid_mutational_signatures_tp_final.RData` | Absolute and fractional SBS2 and SBS13 contributions. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md). Use an output directory for your local results. Available plots are exported as PDF and PNG with Roboto Condensed typography.

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" ExtendedDataFig4
```

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_ExtendedDataFig4(input_dir, output_dir)`.

Shared analysis functions: [age_timing_helpers.R](../functions/age_timing_helpers.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/ExtendedDataFig4.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('## Local figures\n\n')
  files <- file.path(params$figure_dir, c("ExtendedDataFig4a.png","ExtendedDataFig4b.png","ExtendedDataFig4c.png"))
  for (p in files) {
    if (file.exists(p)) {
      cat(sprintf('<img alt="Extended Data Fig. 4: BRAF, TERT, and tumor timing" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    } else {
      cat('Local plot not found: ', basename(p), '\n\n', sep = '')
    }
  }
}
```
