---
title: "Fig. 2: Clonal and subclonal chromosome 22q loss"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](Fig2.R)

## Overview

Copy-number analyses describe the genomic distribution, cellular fraction, and driver associations of chromosome 22q loss in PTC. Clonal and subclonal states are evaluated separately.

## Analyses

### a. Genome-wide copy-number frequencies

Copy-number segments are summarized in 1-Mb bins and plotted as gain and loss frequencies. The module explicitly evaluates the higher-fraction and lower-fraction copy-number states in separate rows. The available implementation provides a two-state frequency view.

### b. Chromosome 22 copy-number heatmap

Segment copy number is expressed relative to a diploid or genome-doubled baseline. The heatmap separates clonal and subclonal states and groups tumors by cohort. Tumors are ordered using the copy-number state of their longest chromosome 22 segment; neutral segments remain visible.

### c. Cellular fraction of subclonal deletion segments

Box plots and individual points summarize the cellular fraction of chromosome 22 deletion segments across cohorts. Each observation is a qualifying segment with a positive subclonal fraction and copy-number loss, so a tumor can contribute multiple observations. This panel does not add statistical-comparison labels.

### d. Associations with driver alterations

Two-sided Fisher exact tests compare driver alteration presence with clonal or subclonal chromosome 22 loss. The available implementation plots log2 odds ratios against −log10 of the unadjusted P value. Gene labels use Holm-adjusted significance within each deletion category. An adjusted-P-value-axis version is not implemented.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData` | Cohort membership and colors. |
| `ngspurity_data.RData` | Battenberg segment states, cellular fractions, purity, and whole-genome-duplication annotations. |
| `Genome_landscape_manual_final.RData` | Curated alterations for association testing. |
| [External GRCh38 cytobands](../data/README.md#genome-reference) | Chromosome and centromere coordinates supplied through `PTC_REFERENCE_DIR` for the genome-wide frequency view. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md). Use an output directory for your local results. Available plots are exported as PDF and PNG with Roboto Condensed typography.

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig2
```

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_Fig2(input_dir, output_dir)`.

Shared analysis functions: [cnv_helpers.R](../functions/cnv_helpers.R), [group_comparison.R](../functions/group_comparison.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig2.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('## Local figures\n\n')
  files <- file.path(params$figure_dir, c("Fig2a_states.png","Fig2b.png","Fig2c.png","Fig2d_raw_p.png"))
  for (p in files) {
    if (file.exists(p)) {
      cat(sprintf('<img alt="Fig. 2: Clonal and subclonal chromosome 22q loss" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    } else {
      cat('Local plot not found: ', basename(p), '\n\n', sep = '')
    }
  }
}
```
