---
title: "Fig. 3: Tumor heterogeneity and subclonal architecture"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](Fig3.R)

## Overview

These analyses concern mutation-level heterogeneity, its relationship to driver alterations, and the detection of subclonal populations. The module provides MATH-score preparation, but the plotting and model implementations for panels a–e are incomplete.

## Analyses

### a–c. MATH scores and driver associations

The preparation function computes mutant-allele tumor heterogeneity (MATH) summaries separately from cancer cell fraction (CCF) and variant allele fraction (VAF). For either variable, the score is `100 × mad(x) / median(x)`, using the default scaling constant in R's `mad()` function. The CCF and VAF summaries are distinct measures and should be labeled accordingly.

The cohort-stratified CCF-based association and distribution analyses are implemented in [Extended Data Fig. 2](ExtendedDataFig2.html). The pooled plots for Fig. 3 are not implemented.

### d. Subclone detectability

A plotting implementation for normalized reads per chromosomal copy (NRPCC) and its detection threshold is not available in this module. Sequencing coverage and purity alone do not specify the threshold calculation.

### e. Driver alterations and subclone number

The module does not implement the driver-specific subclone-number analysis.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData` | Analysis cohort and cohort labels. |
| `DP_info_data.RData` | Mutation-level CCF and VAF values and cluster assignments. |
| `Genome_landscape_manual_final.RData` | Curated driver alterations loaded by the shared preparation functions. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md).

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig3
```

The MATH preparation can be called directly in R:

```{r prepare-math-example, eval=FALSE}
source("functions/ptc_runtime.R")
source("functions/evolution_helpers.R")
source("Rscripts/Fig3.R")
math_scores <- prepare_Fig3_math(Sys.getenv("PTC_INPUT_DIR"))
```

The figure entry point does not currently produce plots.

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_Fig3(input_dir, output_dir)`.

Shared analysis functions: [evolution_helpers.R](../functions/evolution_helpers.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig3.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('No local plot implementation is available for this figure.\n')
}
```
