---
title: "Fig. 4: Mutation-cluster architecture and driver associations"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](Fig4.R)

## Overview

Mutation clusters describe the distribution of cancer cell fractions and the relative mutation burden carried by clonal and subclonal populations. Driver associations are examined using the resulting tumor groups.

## Analyses

### a. Cancer cell fraction distributions

Kernel-density curves compare mutation-cluster CCF distributions across cohorts. Dashed reference lines mark the CCF positions specified in the plotting code.

### b. Mutation burden within clusters

Each point represents a mutation cluster. Position on the horizontal axis gives CCF; vertical position and point size give the fraction of assigned mutations in that cluster. A tumor enters the high-subclonal-mutation-burden group when a nonclonal cluster contains more mutations than the sum assigned to its clonal clusters. Cellular fraction and mutation fraction describe different properties.

### c–d. Driver associations

Fisher exact tests assess associations between each eligible driver category and the high-subclonal-mutation-burden group. Panel c pools cohorts; panel d stratifies by cohort. Plots show log2 odds ratios and unadjusted P values. Holm-adjusted P values are also calculated, while the plotted reference lines use nominal and Bonferroni thresholds.

### e. Evolutionary schematic

A conceptual schematic illustrates alternative relationships among driver events and clonal or subclonal chromosome 22q loss. Branch lengths are illustrative and do not estimate elapsed time.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData` | Cohort membership and colors. |
| `DP_info_data.RData` | Cluster CCF, mutation assignments, clonal labels, and mutation totals. |
| `Genome_landscape_manual_final.RData` | Driver mutations, fusions, and copy-number alteration categories. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md). Use an output directory for your local results. Available plots are exported as PDF and PNG with Roboto Condensed typography.

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig4
```

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_Fig4(input_dir, output_dir)`.

Shared analysis functions: [evolution_helpers.R](../functions/evolution_helpers.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig4.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

### Fig4_schematic.R

```{r code-source-2, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig4_schematic.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('## Local figures\n\n')
  files <- file.path(params$figure_dir, c("Fig4a.png","Fig4b.png","Fig4c.png","Fig4d.png","Fig4e.png"))
  for (p in files) {
    if (file.exists(p)) {
      cat(sprintf('<img alt="Fig. 4: Mutation-cluster architecture and driver associations" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    } else {
      cat('Local plot not found: ', basename(p), '\n\n', sep = '')
    }
  }
}
```
