---
title: "Fig. 6: Clock-like clonal mutations and diagnosis age"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](Fig6.R)

## Overview

Probability-weighted SBS1 and SBS5 contributions are summed over mutations labeled clonal, then related to age at diagnosis. Tumors carrying both BRAF and TERT driver mutations are highlighted.

## Analyses

### a. TCGA-THCA segmented regression

A linear model includes diagnosis age, tumor purity, and sex. Segmented regression estimates an age breakpoint and the slopes before and after it. The prediction curve uses mean tumor purity and the most common sex category. The plot includes a breakpoint confidence interval and uncertainty around the fitted mean.

### b. Chornobyl linear regression

A linear model estimates the age coefficient with adjustment for tumor purity and sex. The displayed smoothing line is a separate, unadjusted regression of mutation burden on age; the annotated slope comes from the adjusted model. These two estimates answer different statistical questions and may differ.

The output places both cohort plots side by side. Clock-like mutation burden is an aggregate signature contribution rather than a direct measurement of tumor age.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData`, `wgs_covdata_tp.RData` | Cohort labels, diagnosis age, tumor purity, and sex. |
| `wgs_clinical.RData`, `Genome_landscape_manual_final.RData` | Clinical data required by the common loader and curated driver groups. |
| `thyroid_signature_probability_tp_final.RData` | Per-mutation SBS1 and SBS5 attribution probabilities. |
| `DP_info_data.RData` | Mutation identifiers and clonal assignments. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md). Use an output directory for your local results. Available plots are exported as PDF and PNG with Roboto Condensed typography.

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig6
```

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_Fig6(input_dir, output_dir)`.

Shared analysis functions: [age_timing_helpers.R](../functions/age_timing_helpers.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig6.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('## Local figures\n\n')
  files <- file.path(params$figure_dir, c("Fig6.png"))
  for (p in files) {
    if (file.exists(p)) {
      cat(sprintf('<img alt="Fig. 6: Clock-like clonal mutations and diagnosis age" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    } else {
      cat('Local plot not found: ', basename(p), '\n\n', sep = '')
    }
  }
}
```
