---
title: "Supplementary Fig. 3: CCF-based heterogeneity by driver group"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](SupplementaryFig3.R)

## Overview

These cohort-stratified analyses summarize the relationship between curated driver alterations and CCF-based MATH scores. They use the shared implementation described in [Extended Data Fig. 2](ExtendedDataFig2.html).

## Analyses

### a. Association tests

MATH is calculated as `100 × mad(CCF) / median(CCF)` within each tumor, using R's default MAD scaling. Wilcoxon tests compare alteration-positive and alteration-negative tumors separately by cohort. The plot shows the log2 ratio of group medians and unadjusted P values; Holm adjustment is calculated within each cohort.

### b. Score distributions

Violin plots with box-plot overlays show MATH distributions by alteration presence and cohort. Alteration categories are selected when a nominal association P value is below 0.05 in any cohort. The plotted outcome is CCF-based MATH.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| `wgsdata_tp.RData` | Cohort membership and colors. |
| `DP_info_data.RData` | Mutation-level CCF and VAF values. |
| `Genome_landscape_manual_final.RData` | Curated alteration categories. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md). Use an output directory for your local results. Available plots are exported as PDF and PNG with Roboto Condensed typography.

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" SupplementaryFig3
```

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_SupplementaryFig3(input_dir, output_dir)`.

Shared analysis functions: [evolution_helpers.R](../functions/evolution_helpers.R).

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/SupplementaryFig3.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

### ExtendedDataFig2.R

```{r code-source-2, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/ExtendedDataFig2.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('## Local figures\n\n')
  files <- file.path(params$figure_dir, c("SupplementaryFig3a.png","SupplementaryFig3b.png"))
  for (p in files) {
    if (file.exists(p)) {
      cat(sprintf('<img alt="Supplementary Fig. 3: CCF-based heterogeneity by driver group" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    } else {
      cat('Local plot not found: ', basename(p), '\n\n', sep = '')
    }
  }
}
```
