---
title: "Supplementary Fig. 1: Allele-specific copy-number profiles"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<style>
body, h1, h2, h3, h4, table {font-family: 'Roboto Condensed', sans-serif;}
body {font-size: 17px; line-height: 1.5;}
pre {font-size: 13px;}
table {font-size: 15px;}
h2 {margin-top: 1.4em;}
</style>

[PTC overview](../README.md) · [Input definitions](../data/input_manifest.tsv) · [Analysis script](SupplementaryFig1.R)

## Overview

These example profiles combine sequencing intensity, allelic imbalance, and inferred copy-number states to illustrate clonal and subclonal copy-number changes.

## Analyses

### a–c. Copy-number profile examples

LogR summarizes relative sequencing intensity, while B-allele frequency (BAF) describes allelic imbalance at informative sites. Battenberg combines these measurements with purity and ploidy estimates to infer allele-specific copy number and cellular fractions.

The sample-profile plotting implementation is not included. Segment summaries alone do not provide the probe-level LogR and BAF tracks required for these displays.

## Inputs

The [input definitions](../data/input_manifest.tsv) describe the expected files and R objects. Participant-level inputs are supplied separately in `PTC_INPUT_DIR`.

| Input | Purpose |
| --- | --- |
| Probe-level LogR and BAF tracks | Required for genome-wide intensity and allelic-imbalance displays. |
| Battenberg segment, purity, and ploidy outputs | Required to overlay inferred copy-number states and cellular fractions. |
| Battenberg plotting inputs | The complete example-plot inputs are not included in this module. |

## Usage

Run from the `PTC` directory after following the [setup instructions](../README.md).

```bash
Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" SupplementaryFig1
```

The figure entry point does not currently produce plots.

This page displays code without running the analysis. Local plot previews are optional through the `include_figures` and `figure_dir` rendering parameters; previews are disabled by default.

## R code

The figure entry point is `run_SupplementaryFig1(input_dir, output_dir)`.

Input loading, plot export, and typography are provided by [ptc_runtime.R](../functions/ptc_runtime.R).

```{r code-source-1, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/SupplementaryFig1.R', warn = FALSE), collapse = '\n'), '\n```\n', sep = '')
```

```{r local-figures, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  cat('No local plot implementation is available for this figure.\n')
}
```
