---
title: "Driver-defined evolutionary routes in papillary thyroid cancer"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
provenance <- read.delim('provenance/figure_provenance.tsv', check.names=FALSE, stringsAsFactors=FALSE)
```

<style>body, h1, h2, h3, h4, table {font-family:'Roboto Condensed',sans-serif;} body {font-size:17px;} pre {font-size:13px;} .table {font-size:15px;} .note {border-left:3px solid #007BBD;padding:12px 16px;background:#f5f9fb;}</style>

## Overview

**Extended Data Fig. 4.** This page follows the UCEC figure-by-figure documentation format.
The source manuscript defines each panel below. Every panel links its Keynote object,
original filename, historical code, required inputs and cleaned implementation.

<div class="note">This is a documentation build. It does not execute the private-data
analysis. Recorded reproduction status comes from the dated local validation audit.
Missing final generators and related historical outputs are explicitly distinguished.</div>

[Package README](../README.md) · [Master provenance](../provenance/figure_provenance.tsv) ·
[Methods requiring review](../provenance/methodology_review.md) · [Input schema](../data/input_manifest.tsv)

## Libraries and execution

Run from the `TCGA-PTC` directory with the required private inputs and Roboto Condensed installed.

```{r execute-example, eval=FALSE}
# Runs this figure in a fresh R process; outputs include per-panel execution status.
# Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" ExtendedDataFig4
```

The executable code appears below and is also available as [ExtendedDataFig4.R](ExtendedDataFig4.R).
Shared functions are in the package's `functions/` directory.

## Extended Data Fig. 4a

a, Estimated age at the most recent common ancestor (MRCA) in tumors classified as other, BRAF only, BRAF plus TERT-promoter mutation or TERT-promoter mutation only.

**Status:** reproduced. <br>
**Original file:** `TP_mrca_group_braf_tert.pdf`. <br>
**Keynote:** slide 11, object `1327078`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 3363-3370;3408-3437. <br>
**Cleaned code:** [ExtendedDataFig4.R](ExtendedDataFig4.R) (`Rscripts/ExtendedDataFig4.R:3 (run_ExtendedDataFig4)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_evolution_analysis/Chronological_timing_short.RData`.

Group sizes, boxes, distribution and significant comparisons visually match. Current rstatix rounding yields4.87e-6 vs original4.86e-6 and3.37e-4 vs3.38e-4. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-ExtendedDataFig4-a, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("ExtendedDataFig4a.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Extended Data Fig. 4b

b, Age at diagnosis across the same groups. In a,b, points represent individual tumors and box plots show the median (center line), interquartile range (box) and whiskers extending to 1.5× the interquartile range; FDR-adjusted P values for the indicated pairwise comparisons are shown.

**Status:** reproduced. <br>
**Original file:** `age_tert_braf.pdf`. <br>
**Keynote:** slide 11, object `1327077`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 3007-3038. <br>
**Cleaned code:** [ExtendedDataFig4.R](ExtendedDataFig4.R) (`Rscripts/ExtendedDataFig4.R:3 (run_ExtendedDataFig4)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData`.

Groups, boxes and significant comparisons match; one displayed BH P differs at final rounding digit4.77e-5 vs4.78e-5. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-ExtendedDataFig4-b, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("ExtendedDataFig4b.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Extended Data Fig. 4c

c, Prevalence of APOBEC mutagenesis across TERT-only, other, BRAF-only and BRAF plus TERT tumors. APOBEC mutagenesis was detected in 9.1%, 11.9%, 21.8% and 39.0% of tumors, respectively.

**Status:** reproduced. <br>
**Original file:** `tp_apobec_braf_tert.pdf`. <br>
**Keynote:** slide 11, object `1327076`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 1640-1689. <br>
**Cleaned code:** [ExtendedDataFig4.R](ExtendedDataFig4.R) (`Rscripts/ExtendedDataFig4.R:3 (run_ExtendedDataFig4)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_mutational_signatures_tp_final.RData`.

All eight labeled proportions match: present9.1/11.9/21.8/39.0%; absent90.9/88.1/78.2/61.0%. Strict50/5% thresholds retained. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-ExtendedDataFig4-c, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("ExtendedDataFig4c.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Cleaned executable source

The code displayed here is read from the executable `.R` file at render time,
so the report and source cannot silently diverge.

```{r code-source, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/ExtendedDataFig4.R', warn=FALSE), collapse='\n'), '\n```\n', sep='')
```

## Validation and limitations

```{r validation-table, echo=FALSE}
z <- provenance[provenance$figure_id == 'Extended Data Fig. 4', c('manuscript_id','reproducibility_status','local_original_match')]
knitr::kable(z, col.names=c('Panel','Reproduction status','Embedded versus archived original'))
```

A byte-identical archived original confirms artifact identity; it does not by itself
validate cleaned code. Scientific/version discrepancies are documented in the
[methodology review](../provenance/methodology_review.md).

## Documentation build environment

```{r documentation-session, echo=FALSE}
cat(R.version.string)
```
