---
title: "Driver-defined evolutionary routes in papillary thyroid cancer"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
provenance <- read.delim('provenance/figure_provenance.tsv', check.names=FALSE, stringsAsFactors=FALSE)
```

<style>body, h1, h2, h3, h4, table {font-family:'Roboto Condensed',sans-serif;} body {font-size:17px;} pre {font-size:13px;} .table {font-size:15px;} .note {border-left:3px solid #007BBD;padding:12px 16px;background:#f5f9fb;}</style>

## Overview

**Fig. 1.** This page follows the UCEC figure-by-figure documentation format.
The source manuscript defines each panel below. Every panel links its Keynote object,
original filename, historical code, required inputs and cleaned implementation.

<div class="note">This is a documentation build. It does not execute the private-data
analysis. Recorded reproduction status comes from the dated local validation audit.
Missing final generators and related historical outputs are explicitly distinguished.</div>

[Package README](../README.md) · [Master provenance](../provenance/figure_provenance.tsv) ·
[Methods requiring review](../provenance/methodology_review.md) · [Input schema](../data/input_manifest.tsv)

## Libraries and execution

Run from the `TCGA-PTC` directory with the required private inputs and Roboto Condensed installed.

```{r execute-example, eval=FALSE}
# Runs this figure in a fresh R process; outputs include per-panel execution status.
# Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig1
```

The executable code appears below and is also available as [Fig1.R](Fig1.R).
Shared functions are in the package's `functions/` directory.

## Fig. 1a

a, Age at diagnosis, tumor sequencing depth and tumor purity across TCGA-THCA (n = 455), Chornobyl-unexposed (n = 70) and Chornobyl-exposed (n = 283) tumors. Box plots show the median (center line), interquartile range (box) and whiskers extending to 1.5× the interquartile range.

**Status:** source code partially identified. <br>
**Original file:** `study2_overview.pdf`. <br>
**Keynote:** slide 2, object `1293023`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 60-90 age;97-117 sequencing depth;126-145 purity. <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `wgsdata_tp.RData;wgs_clinical.RData;wgs_metrics.RData;ngspurity_data.RData`.

Exact Keynote filename from slide 2; component relationship follows manuscript legend, not an exact export match. No export or variable trace for study2_overview.pdf in any biowulf R/Rmd. Related component code contains both tumor and normal depth; final composite requires layout/cropping evidence. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-a, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1a_related_age.png","Fig1a_related_depth.png","Fig1a_related_purity.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1b

b, Cohort-specific mutation frequencies of driver genes identified by IntOGen, coloured by predicted mode of action (activating, ambiguous or loss of function).

**Status:** reproduced. <br>
**Original file:** `intogene_drivers3.pdf`. <br>
**Keynote:** slide 2, object `1283038`; entire imported object. <br>
**Historical source:** `Intogene_drivers.R` — 6;58-126;128. <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `intogene_drivers.RData;Genome_landscape_manual_final.RData;wgsdata_tp.RData;wgs_covdata_tp.RData`.

Exact ggsave line126. Minimal replay should explicitly load wgs_covdata snapshot; do not need historical thyroid_maf_freq plots at12-52. Original drivers.tsv mounted source unavailable; saved intogene_drivers.RData contains intogene and can supply the same object. wgs_covdata is referenced but never explicitly loaded by this script. Full ancestry/QC derivation remains upstream. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-b, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1b.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1c

c, Structural-variant breakpoint density per sample in 5-Mb genomic windows across the three cohorts. Recurrent peaks overlapping major kinase genes are labelled.

**Status:** source code partially identified. <br>
**Original file:** `ALL_Tumor_Recurrent_p1_combined.pdf`. <br>
**Keynote:** slide 2, object `1283987`; entire imported object. <br>
**Historical source:** `Recurrent_SV_Plot.R` — 13-21;46-117;129. <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `wgsdata_tp.RData;1170_manta_meerkat_union_window50_highquality.txt;Homo_sapiens_assembly38.fasta.fai`.

Dynamic exact per-cohort filename expression paste0(studyid,"_Recurrent_p1.pdf") verified line129; no expression creates the final _combined file. Combined export and gene hotspot label code absent. Code fixes studyid=names(study_color2)[3]; historical per-cohort filenames require explicitly iterating cohort values. Combined manuscript panel cannot be claimed from absent assembly code. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-c, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1c_related_TCGA_THCA.png","Fig1c_related_Chornobyl_Unexposed.png","Fig1c_related_Chornobyl_Exposed.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1d

d–f, Circos plots showing structural variants in TCGA-THCA (d), Chornobyl-unexposed (e) and Chornobyl-exposed (f) tumors.

**Status:** reproduced. <br>
**Original file:** `SV_circos_plot_thca.pdf`. <br>
**Keynote:** slide 2, object `1332655`; entire imported object. <br>
**Historical source:** `Meerkat/SV_circosplot.R` — 4-21;23;28-40 (export 34, currently commented). <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `svdata.RData;wgsdata_tp.RData`.

Exact filename and matching cohort filter are in source, but neither is active in the saved script. Cohort selection and matching pdf export are commented. Sequential script emits all-cohort file only; explicit paired cohort/output parameterization required. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-d, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1d.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1e

d–f, Circos plots showing structural variants in TCGA-THCA (d), Chornobyl-unexposed (e) and Chornobyl-exposed (f) tumors.

**Status:** reproduced. <br>
**Original file:** `SV_circos_plot_unexposed.pdf`. <br>
**Keynote:** slide 2, object `1332654`; entire imported object. <br>
**Historical source:** `Meerkat/SV_circosplot.R` — 4-21;24;28-40 (export 35, currently commented). <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `svdata.RData;wgsdata_tp.RData`.

Exact filename and matching cohort filter are in source, but neither is active in the saved script. Cohort selection and matching pdf export are commented. Sequential script emits all-cohort file only; explicit paired cohort/output parameterization required. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-e, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1e.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1f

d–f, Circos plots showing structural variants in TCGA-THCA (d), Chornobyl-unexposed (e) and Chornobyl-exposed (f) tumors.

**Status:** reproduced. <br>
**Original file:** `SV_circos_plot_exposed.pdf`. <br>
**Keynote:** slide 2, object `1332658`; entire imported object. <br>
**Historical source:** `Meerkat/SV_circosplot.R` — 4-21;25;28-40 (export 36, currently commented). <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `svdata.RData;wgsdata_tp.RData`.

Exact filename and matching cohort filter are in source, but neither is active in the saved script. Cohort selection and matching pdf export are commented. Sequential script emits all-cohort file only; explicit paired cohort/output parameterization required. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-f, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1f.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1g

g–i, Integrated genomic profiles of driver mutations, oncogenic fusions, TERT-promoter mutations and clonal or subclonal 22q loss in TCGA-THCA (g), Chornobyl-exposed (h) and Chornobyl-unexposed (i) tumors. Bars above each oncoprint indicate sample-level mutation burden; percentages at right indicate cohort-specific frequencies of individual alterations.

**Status:** reproduced. <br>
**Original file:** `Genome_landscape_tcga.pdf`. <br>
**Keynote:** slide 2, object `1284253`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2777-2785;2846-2870;2872-2878;2895-2913. <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `Genome_landscape_manual_final.RData;wgsdata_tp.RData;oncoplot_colors.csv`.

Saved final RData includes data_tmb0, avoiding rebuild of mutation burdens from raw variants. Cohort legend order for g/h/i differs from d/e/f. All three if(TRUE) blocks execute, leaving only exposed state. Three successive save_plot calls overwrite same filename at different dimensions. Must select cohort intentionally and verify intended manuscript dimensions. Underlying manual curation cannot be reconstructed entirely from code alone. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-g, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1g.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1h

g–i, Integrated genomic profiles of driver mutations, oncogenic fusions, TERT-promoter mutations and clonal or subclonal 22q loss in TCGA-THCA (g), Chornobyl-exposed (h) and Chornobyl-unexposed (i) tumors. Bars above each oncoprint indicate sample-level mutation burden; percentages at right indicate cohort-specific frequencies of individual alterations.

**Status:** reproduced. <br>
**Original file:** `Genome_landscape_exposed.pdf`. <br>
**Keynote:** slide 2, object `1284252`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2777-2785;2846-2870;2887-2893;2895-2913. <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `Genome_landscape_manual_final.RData;wgsdata_tp.RData;oncoplot_colors.csv`.

Saved final RData includes data_tmb0, avoiding rebuild of mutation burdens from raw variants. Cohort legend order for g/h/i differs from d/e/f. All three if(TRUE) blocks execute, leaving only exposed state. Three successive save_plot calls overwrite same filename at different dimensions. Must select cohort intentionally and verify intended manuscript dimensions. Underlying manual curation cannot be reconstructed entirely from code alone. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-h, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1h.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 1i

g–i, Integrated genomic profiles of driver mutations, oncogenic fusions, TERT-promoter mutations and clonal or subclonal 22q loss in TCGA-THCA (g), Chornobyl-exposed (h) and Chornobyl-unexposed (i) tumors. Bars above each oncoprint indicate sample-level mutation burden; percentages at right indicate cohort-specific frequencies of individual alterations.

**Status:** reproduced. <br>
**Original file:** `Genome_landscape_chornobyl_unexposed.pdf`. <br>
**Keynote:** slide 2, object `1284251`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2777-2785;2846-2870;2879-2885;2895-2913. <br>
**Cleaned code:** [Fig1.R](Fig1.R) (`Rscripts/Fig1.R:4 (run_Fig1)`).

**Inputs:** `Genome_landscape_manual_final.RData;wgsdata_tp.RData;oncoplot_colors.csv`.

Saved final RData includes data_tmb0, avoiding rebuild of mutation burdens from raw variants. Cohort legend order for g/h/i differs from d/e/f. All three if(TRUE) blocks execute, leaving only exposed state. Three successive save_plot calls overwrite same filename at different dimensions. Must select cohort intentionally and verify intended manuscript dimensions. Underlying manual curation cannot be reconstructed entirely from code alone. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig1-i, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig1i.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Cleaned executable source

The code displayed here is read from the executable `.R` file at render time,
so the report and source cannot silently diverge.

```{r code-source, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig1.R', warn=FALSE), collapse='\n'), '\n```\n', sep='')
```

## Validation and limitations

```{r validation-table, echo=FALSE}
z <- provenance[provenance$figure_id == 'Fig. 1', c('manuscript_id','reproducibility_status','local_original_match')]
knitr::kable(z, col.names=c('Panel','Reproduction status','Embedded versus archived original'))
```

A byte-identical archived original confirms artifact identity; it does not by itself
validate cleaned code. Scientific/version discrepancies are documented in the
[methodology review](../provenance/methodology_review.md).

## Documentation build environment

```{r documentation-session, echo=FALSE}
cat(R.version.string)
```
