---
title: "Driver-defined evolutionary routes in papillary thyroid cancer"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
provenance <- read.delim('provenance/figure_provenance.tsv', check.names=FALSE, stringsAsFactors=FALSE)
```

<style>body, h1, h2, h3, h4, table {font-family:'Roboto Condensed',sans-serif;} body {font-size:17px;} pre {font-size:13px;} .table {font-size:15px;} .note {border-left:3px solid #007BBD;padding:12px 16px;background:#f5f9fb;}</style>

## Overview

**Fig. 3.** This page follows the UCEC figure-by-figure documentation format.
The source manuscript defines each panel below. Every panel links its Keynote object,
original filename, historical code, required inputs and cleaned implementation.

<div class="note">This is a documentation build. It does not execute the private-data
analysis. Recorded reproduction status comes from the dated local validation audit.
Missing final generators and related historical outputs are explicitly distinguished.</div>

[Package README](../README.md) · [Master provenance](../provenance/figure_provenance.tsv) ·
[Methods requiring review](../provenance/methodology_review.md) · [Input schema](../data/input_manifest.tsv)

## Libraries and execution

Run from the `TCGA-PTC` directory with the required private inputs and Roboto Condensed installed.

```{r execute-example, eval=FALSE}
# Runs this figure in a fresh R process; outputs include per-panel execution status.
# Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig3
```

The executable code appears below and is also available as [Fig3.R](Fig3.R).
Shared functions are in the package's `functions/` directory.

## Fig. 3a

a, Ridgeline distributions of MATH scores across the three cohorts; vertical lines indicate cohort medians.

**Status:** source code partially identified. <br>
**Original file:** `tp_tumor_hetrogenerity_math_ccf2.pdf`. <br>
**Keynote:** slide 4, object `1303194`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2366-2410. <br>
**Cleaned code:** [Fig3.R](Fig3.R) (`Rscripts/Fig3.R:8 (run_Fig3)`).

**Inputs:** `wgsdata_tp.RData;DP_info_data.RData`.

Exact final filename and ridgeline plot code absent; related historical export is box/point comparison. Manuscript describes VAF dispersion but historical plotted object is MATH_CCF. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

## Fig. 3b

b, Associations of recurrent genomic alterations with MATH score in the combined cohort. The x axis shows the log2 fold change in MATH between alteration-positive and alteration-negative tumors, and the y axis shows −log10(FDR). The horizontal dotted line denotes FDR = 0.05.

**Status:** source code partially identified. <br>
**Original file:** `tp_tumor_hetrogenerity_math_ccf_drivers_all.pdf`. <br>
**Keynote:** slide 4, object `1305162`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2366-2387;2461-2513. <br>
**Cleaned code:** [Fig3.R](Fig3.R) (`Rscripts/Fig3.R:8 (run_Fig3)`).

**Inputs:** `wgsdata_tp.RData;DP_info_data.RData;Genome_landscape_manual_final.RData`.

Exact combined-cohort filename and generating code absent. Historical code is stratified by Study2 and uses raw-P y-axis; cannot relabel it as combined-FDR panel. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

## Fig. 3c

c, MATH score distributions stratified by TERT-promoter mutation, BRAF mutation, KRAS mutation and clonal 22q-loss status.

**Status:** source code partially identified. <br>
**Original file:** `tp_tumor_hetrogenerity_math_ccf_drivers2_all.pdf`. <br>
**Keynote:** slide 4, object `1305537`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2366-2387;2461-2490;2520-2534. <br>
**Cleaned code:** [Fig3.R](Fig3.R) (`Rscripts/Fig3.R:8 (run_Fig3)`).

**Inputs:** `wgsdata_tp.RData;DP_info_data.RData;Genome_landscape_manual_final.RData`.

Exact combined-cohort filename and generating code absent. Current historical plot facets by Study2; manuscript panel combines cohorts. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

## Fig. 3d

d, Minimum CCF of detected mutation clusters as a function of the number of reads per tumor chromosomal copy (NRPCC). Each point represents one tumor and colour indicates the number of inferred subclones; marginal CCF distributions are shown at right.

**Status:** unresolved. <br>
**Original file:** `NRPCC_threshold.pdf`. <br>
**Keynote:** slide 4, object `1285819`; entire imported object. <br>
**Historical source:** `unresolved` — . <br>
**Cleaned code:** [Fig3.R](Fig3.R) (`Rscripts/Fig3.R:8 (run_Fig3)`).

**Inputs:** `Potential inputs only: ngspurity_data.RData;wgs_covdata_tp.RData;DP_info_data.RData`.

Input suggestions are conceptual candidates, not a recovered dependency graph. Do not invent NRPCC or minimum-detection methodology. No NRPCC, NRPCC_threshold, min_ccf, or minimum-CCF plotting logic found in any biowulf R/Rmd. Formula, threshold, cluster filtering and plotting code unverified. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

## Fig. 3e

e, Distribution of the number of detected subclones across major genomic alteration groups. Counts are shown within bars.

**Status:** source code partially identified. <br>
**Original file:** `drivers_number_subclones.pdf`. <br>
**Keynote:** slide 4, object `1285891`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2115-2152;2461-2465 related only. <br>
**Cleaned code:** [Fig3.R](Fig3.R) (`Rscripts/Fig3.R:8 (run_Fig3)`).

**Inputs:** `ngspurity_data.RData;wgsdata_tp.RData;Genome_landscape_manual_final.RData`.

Exact filename and driver-stratified subclone-count plotting logic absent. Manuscript >=50 mutation cluster threshold is not enforced in related count block. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

## Cleaned executable source

The code displayed here is read from the executable `.R` file at render time,
so the report and source cannot silently diverge.

```{r code-source, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig3.R', warn=FALSE), collapse='\n'), '\n```\n', sep='')
```

## Validation and limitations

```{r validation-table, echo=FALSE}
z <- provenance[provenance$figure_id == 'Fig. 3', c('manuscript_id','reproducibility_status','local_original_match')]
knitr::kable(z, col.names=c('Panel','Reproduction status','Embedded versus archived original'))
```

A byte-identical archived original confirms artifact identity; it does not by itself
validate cleaned code. Scientific/version discrepancies are documented in the
[methodology review](../provenance/methodology_review.md).

## Documentation build environment

```{r documentation-session, echo=FALSE}
cat(R.version.string)
```
