---
title: "Driver-defined evolutionary routes in papillary thyroid cancer"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
provenance <- read.delim('provenance/figure_provenance.tsv', check.names=FALSE, stringsAsFactors=FALSE)
```

<style>body, h1, h2, h3, h4, table {font-family:'Roboto Condensed',sans-serif;} body {font-size:17px;} pre {font-size:13px;} .table {font-size:15px;} .note {border-left:3px solid #007BBD;padding:12px 16px;background:#f5f9fb;}</style>

## Overview

**Fig. 5.** This page follows the UCEC figure-by-figure documentation format.
The source manuscript defines each panel below. Every panel links its Keynote object,
original filename, historical code, required inputs and cleaned implementation.

<div class="note">This is a documentation build. It does not execute the private-data
analysis. Recorded reproduction status comes from the dated local validation audit.
Missing final generators and related historical outputs are explicitly distinguished.</div>

[Package README](../README.md) · [Master provenance](../provenance/figure_provenance.tsv) ·
[Methods requiring review](../provenance/methodology_review.md) · [Input schema](../data/input_manifest.tsv)

## Libraries and execution

Run from the `TCGA-PTC` directory with the required private inputs and Roboto Condensed installed.

```{r execute-example, eval=FALSE}
# Runs this figure in a fresh R process; outputs include per-panel execution status.
# Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig5
```

The executable code appears below and is also available as [Fig5.R](Fig5.R).
Shared functions are in the package's `functions/` directory.

## Fig. 5a

a, Associations of recurrent genomic alterations with age at diagnosis, age at exposure and radiation dose. The x axis shows the effect estimate and the y axis shows −log10(FDR); symbol shape denotes alteration type and colour denotes the age or exposure variable.

**Status:** reproduced. <br>
**Original file:** `TP_genomiclandscape_associaiton.pdf`. <br>
**Keynote:** slide 6, object `1284819`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 3046-3058;3158;3178-3215. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData`.

Original-vs-cleaned logistic coefficients, SE, P and BH P match exactly; visual data geometry matches. Matrix excludes tumors with no curated alterations. Two original/cleaned glm fits warn of fitted probabilities 0/1. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig5-a, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig5a.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 5b

b–d, Age-stratified frequencies of TERT-promoter mutation (b), BRAF mutation (c) and oncogenic fusions (d) by cohort. Point size indicates the number of tumors in each age group.

**Status:** reproduced. <br>
**Original file:** `age_tert.pdf`. <br>
**Keynote:** slide 6, object `1284893`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2920-2927;2979-3001. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData`.

Cohort/age point frequencies visually match Keynote source; age-bin 70 retained in 60-70. TERT any event. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig5-b, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig5b.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 5c

b–d, Age-stratified frequencies of TERT-promoter mutation (b), BRAF mutation (c) and oncogenic fusions (d) by cohort. Point size indicates the number of tumors in each age group.

**Status:** reproduced. <br>
**Original file:** `age_braf.pdf`. <br>
**Keynote:** slide 6, object `1284891`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2920-2950. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData`.

Cohort/age point frequencies visually match; BRAF Mutation_Driver only. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig5-c, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig5c.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 5d

b–d, Age-stratified frequencies of TERT-promoter mutation (b), BRAF mutation (c) and oncogenic fusions (d) by cohort. Point size indicates the number of tumors in each age group.

**Status:** reproduced. <br>
**Original file:** `age_fusion.pdf`. <br>
**Keynote:** slide 6, object `1284892`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 2920-2927;2954-2975. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData`.

Cohort/age point frequencies visually match; any Fusion membership. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig5-d, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig5d.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 5e

e, Associations of clinical, ancestry and exposure variables with inferred tumor latency combined cohorts. Points and error bars show regression coefficients and confidence intervals, respectively; β coefficients and P values are shown.

**Status:** reproduced. <br>
**Original file:** `TP_Latency_associaiton_exposure.pdf`. <br>
**Keynote:** slide 6, object `1310742`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 3363-3370;3453-3489. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_evolution_analysis/Chronological_timing_short.RData`.

Embedded Keynote PDF visually compared. All six coefficients/P values match; numerical maximum delta8.53e-14. Local same-name PDF is a different annotation version. 1x intermediate, no upstream timing rerun. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig5-e, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig5e.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 5f

f, Associations of recurrent genomic alterations with inferred tumor latency.

**Status:** reproduced. <br>
**Original file:** `TP_latency_drivergenes_association.pdf`. <br>
**Keynote:** slide 6, object `1310743`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 3363-3370;3495-3553. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_evolution_analysis/Chronological_timing_short.RData`.

Embedded Keynote PDF visually compared. Original coefficients/SE/P/Holm-adjusted P match exactly. Axis corrected explicitly to Holm-adjusted P; original says FDR. Local same-name PDF differs in annotation version. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

```{r preview-Fig5-f, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig5f.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 5g

g, Associations of recurrent genomic alterations with inferred age at the most recent common ancestor (MRCA). In f and g, the x axis shows the regression coefficient and the y axis shows −log10(FDR). Dashed lines denote FDR = 0.05.

**Status:** source code partially identified. <br>
**Original file:** `TP_MRCA_drivergenes_association.pdf`. <br>
**Keynote:** slide 6, object `1310744`; entire imported object. <br>
**Historical source:** `TP_Comparision.R` — 3495-3514. <br>
**Cleaned code:** [Fig5.R](Fig5.R) (`Rscripts/Fig5.R:5 (run_Fig5)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_evolution_analysis/Chronological_timing_short.RData`.

Exact filename recovered and local PDF equals Keynote original. Nearby latency model shares preparation, but no exact MRCA model/export found. No response-swapped speculative model supplied. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF.

## Cleaned executable source

The code displayed here is read from the executable `.R` file at render time,
so the report and source cannot silently diverge.

```{r code-source, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig5.R', warn=FALSE), collapse='\n'), '\n```\n', sep='')
```

## Validation and limitations

```{r validation-table, echo=FALSE}
z <- provenance[provenance$figure_id == 'Fig. 5', c('manuscript_id','reproducibility_status','local_original_match')]
knitr::kable(z, col.names=c('Panel','Reproduction status','Embedded versus archived original'))
```

A byte-identical archived original confirms artifact identity; it does not by itself
validate cleaned code. Scientific/version discrepancies are documented in the
[methodology review](../provenance/methodology_review.md).

## Documentation build environment

```{r documentation-session, echo=FALSE}
cat(R.version.string)
```
