---
title: "Driver-defined evolutionary routes in papillary thyroid cancer"
author: "Tongwu Zhang"
date: "2026-09-15"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    self_contained: true
params:
  include_figures: false
  figure_dir: "Rscripts/Figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
provenance <- read.delim('provenance/figure_provenance.tsv', check.names=FALSE, stringsAsFactors=FALSE)
```

<style>body, h1, h2, h3, h4, table {font-family:'Roboto Condensed',sans-serif;} body {font-size:17px;} pre {font-size:13px;} .table {font-size:15px;} .note {border-left:3px solid #007BBD;padding:12px 16px;background:#f5f9fb;}</style>

## Overview

**Fig. 6.** This page follows the UCEC figure-by-figure documentation format.
The source manuscript defines each panel below. Every panel links its Keynote object,
original filename, historical code, required inputs and cleaned implementation.

<div class="note">This is a documentation build. It does not execute the private-data
analysis. Recorded reproduction status comes from the dated local validation audit.
Missing final generators and related historical outputs are explicitly distinguished.</div>

[Package README](../README.md) · [Master provenance](../provenance/figure_provenance.tsv) ·
[Methods requiring review](../provenance/methodology_review.md) · [Input schema](../data/input_manifest.tsv)

## Libraries and execution

Run from the `TCGA-PTC` directory with the required private inputs and Roboto Condensed installed.

```{r execute-example, eval=FALSE}
# Runs this figure in a fresh R process; outputs include per-panel execution status.
# Rscript Rscripts/run_all.R "$PTC_INPUT_DIR" "$PTC_OUTPUT_DIR" Fig6
```

The executable code appears below and is also available as [Fig6.R](Fig6.R).
Shared functions are in the package's `functions/` directory.

## Fig. 6a

a, Clonal SBS1 plus SBS5 mutation burden as a function of age at diagnosis in TCGA-THCA. Segmented regression identified a breakpoint at 52.2 years (95% confidence interval, 45.8–58.7 years), with age-associated slopes of 4.7 ± 1.5 mutations per year before and 15.8 ± 1.8 mutations per year after the breakpoint. Red-inlined points indicate tumors harboring TERT-promoter mutations.

**Status:** reproduced. <br>
**Original file:** `tp_age_clock_mutation_acc.pdf`. <br>
**Keynote:** slide 7, object `1288144`; left half. <br>
**Historical source:** `Segmented_breakpoint_analysis.R;biowulf/TP_Comparision.R` — 10-113;1697-1747. <br>
**Cleaned code:** [Fig6.R](Fig6.R) (`Rscripts/Fig6.R:4 (run_Fig6)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_signature_probability_tp_final.RData;DP_info_data.RData`.

Both panels share composite imported PDF. Numerical breakpoint52.2444 and printed95%CI45.8-58.7/slopes4.7,15.8,4.6 match Keynote. Data distributions and highlights visually match. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF. The same composite PDF object contains both panels: a is left TCGA segmented regression; b is right Chornobyl linear regression. Author confirmed in this session that both panels came from one file and panel labels a/b were added manually in Keynote; this is intentional, not a mapping conflict.

```{r preview-Fig6-a, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig6.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Fig. 6b

b, Corresponding relationship in the combined Chornobyl cohorts, with an age-associated linear slope of 4.6 mutations per year. Points represent individual tumors and shaded bands indicate uncertainty around the fitted regression lines.

**Status:** reproduced. <br>
**Original file:** `tp_age_clock_mutation_acc.pdf`. <br>
**Keynote:** slide 7, object `1288144`; right half. <br>
**Historical source:** `Segmented_breakpoint_analysis.R;biowulf/TP_Comparision.R` — 10-113;1697-1747. <br>
**Cleaned code:** [Fig6.R](Fig6.R) (`Rscripts/Fig6.R:4 (run_Fig6)`).

**Inputs:** `wgsdata_tp.RData;wgs_covdata_tp.RData;wgs_clinical.RData;Genome_landscape_manual_final.RData;thyroid_signature_probability_tp_final.RData;DP_info_data.RData`.

Same composite source as Fig6a. Chornobyl drawn lm line unadjusted, slope label adjusted; highlight actually BRAF+TERT intersection. These source discrepancies retained; no claim of corrected methodology. Imported original PDF filename recovered verbatim from Metadata.iwa preferred_file_name; visible label and plot content confirmed in native Keynote PDF. The same composite PDF object contains both panels: a is left TCGA segmented regression; b is right Chornobyl linear regression. Author confirmed in this session that both panels came from one file and panel labels a/b were added manually in Keynote; this is intentional, not a mapping conflict.

```{r preview-Fig6-b, echo=FALSE, results='asis'}
if (isTRUE(params$include_figures)) {
  files <- file.path(params$figure_dir, c("Fig6.png"))
  for (p in files) {
    if (file.exists(p)) cat(sprintf('<img alt="Local validation plot" src="%s" style="max-width:100%%;height:auto;" />', knitr::image_uri(p)))
    else cat('Expected local plot is absent: ', basename(p), '\n\n')
  }
}
```

## Cleaned executable source

The code displayed here is read from the executable `.R` file at render time,
so the report and source cannot silently diverge.

```{r code-source, echo=FALSE, results='asis'}
cat('```r\n', paste(readLines('Rscripts/Fig6.R', warn=FALSE), collapse='\n'), '\n```\n', sep='')
```

## Validation and limitations

```{r validation-table, echo=FALSE}
z <- provenance[provenance$figure_id == 'Fig. 6', c('manuscript_id','reproducibility_status','local_original_match')]
knitr::kable(z, col.names=c('Panel','Reproduction status','Embedded versus archived original'))
```

A byte-identical archived original confirms artifact identity; it does not by itself
validate cleaned code. Scientific/version discrepancies are documented in the
[methodology review](../provenance/methodology_review.md).

## Documentation build environment

```{r documentation-session, echo=FALSE}
cat(R.version.string)
```
